Wetting of the adhesive fluid controls underwater adhesion in insects

### 2021-02-16 source/tex/ladybug_adhesion.tex
moved appendix section to separate supplementary file.
hard coded references pointing to supplementary file.

%% LyX 2.3.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{achemso}
\usepackage[T2A,LGR,T1]{fontenc}
\usepackage{babel}
\usepackage{array}
\usepackage{textcomp}
\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subscript}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}

@@

@@

In air, adhesion forces of the distal pad of ladybug beetle against
glass and PFOTS were similar, i.e. no significant differences were
detected (Figure \ref{fig:Effect-of-contact} and Table \ref{tab:Statistical-analysis}).
In contrast, the underwater adhesion on a PFOTS surface was significantly
larger than on glass (p < 0.001). This stronger adhesion on PFOTS
was observed both in the presence and absence of a trapped bubble.
In both cases, the adhesion force reached similar values as in air. In contrast,
on glass, adhesion underwater was significantly reduced when compared
to dry conditions, irrespective of the presence of a trapped bubble
(p $\leq$ 0.002). In the presence of a bubble, underwater adhesion on glass was slightly higher (CLES = 0.84, p = 0.07).

@@

The capillary force, $f$, is the sum of two contributions: surface
tension and Laplace pressure (equation \ref{eq:f_bridge}). Force
versus distance for a single capillary bridge was calculated by Surface
Evolver simulations\cite{RN206}  (described in \ref{subsec:Simulation-Method}),
and used to obtain $F_{net}$ as a function of $d$ for each type
of contact. The adhesion force of the complete hairy pad system was
then obtained from the minima of $F_{net}$, where negative force
values represent attraction.

@@

\subcaption{Force contributions}%
\end{minipage}\caption{\label{fig:Single-bridge}Normalised capillary force of a single liquid
bridge in contact with a substrate and pinned to a circular perimeter
on top. Fluid size parameter, $\phi_{f}=2$. Negative force values
represents attraction. a) Force-distance curves are shown for different
contact angles of the liquid with the substrate. b) Adhesion forces,
calculated from the minima of the corresponding force-distance curves,
are plotted as a function of contact angle with the substrate, together
with its Laplace and surface tension components (equation \ref{eq:f_bridge}).
Simulation snapshots of the liquid meniscus corresponding to angles
6\textdegree{} and 150\textdegree{} are depicted.}
\end{figure}

@@

Our experiments demonstrate that the ladybug beetle can attach underwater
to a hydrophobic substrate even without a bubble trapped around its
hairs. A previous study\cite{RN87} proposed that a bubble
is necessary for underwater attachment in beetles. This is, however,
only true for hydrophilic substrates, where a trapped bubble can facilitate
underwater adhesion due to the hairs making contact in a de-wetted
environment. For a hydrophobic substrate, the adhesion is similar
regardless of whether the contact occurs in air or underwater conditions,
with or without a trapped bubble. Our theoretical calculations further
show that the bubble by itself has a negligible capillary contribution
to the net underwater adhesion of the pad. Direct force measurement
of a single similarly sized bubble making contact with a hydrophobic
substrate shows a maximum adhesion less than 50 \textmu N, which
further validates that the bubble's contribution is insignificant
(\ref{subsec:Capillary-force-due}).

@@

We are grateful to Eduard Arzt and Ren\textcyr{\`\cyre} Hensel for
fruitful discussions. This work was funded by the \emph{Deutsche Forschungsgemeinschaft}
(Grant number: PI 1351/2-1).

\appendix

\section{Appendix}
%reset figure/table/equation numbers and include prefix
\setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}} 
\setcounter{table}{0} \renewcommand{\thetable}{A.\arabic{table}} 
\setcounter{equation}{0} \renewcommand{\theequation}{A.\arabic{equation}}

\subsection{Simulation method: Single capillary bridge \label{subsec:Simulation-Method}}

 Capillary force due to a single adhesive fluid or bubble meniscus (termed ``capillary
bridge'') is calculated by performing simulations in Surface Evolver
\cite{RN206}, similar to the method described by \citet{RN93}.
A simple cubic geometry, mimicking the capillary bridge, of constant
volume, $V$, is defined as the initial condition with an interfacial
tension, $\gamma$, with the surrounding medium. Interfacial tension
of the capillary bridge with the substrate is given by $\gamma\cos\theta$,
where $\theta$ is the corresponding contact angle inside the bridge.
For the case of a bubble meniscus, $\theta$ is defined w.r.t. the
surrounding water, since $\theta$ can also directly characterise
the substrate wettability. The capillary bridge spans a gap distance
$d$ between the top face and the substrate. The boundary conditions
are set corresponding to a pinned contact line of diameter $D$ on
the top face and constant interfacial tension with the substrate on
the bottom. All lengths are normalised relative to length $s=\left(3V/4\pi\right)^{1/3}$.
An appropriate geometry refinement routine is chosen to evolve the
capillary bridge shape to its minimum energy state. The normalised
total capillary force, $\hat{f}=f/\gamma s$, is the sum of the Laplace
pressure and surface tension contributions , where:

\begin{equation}
f=f_{laplace}+f_{surface\,tension}=\varDelta P_{laplace}A_{_{bottom}}+2\pi R_{bottom}\gamma\sin\theta\label{eq:f_bridge}
\end{equation}

Here, $\varDelta P_{laplace}$ is the Laplace pressure of the equilibrium
capillary bridge, $A_{bottom}$ is the contact area of the capillary
bridge with the substrate at bottom and $R_{bottom}$ is the corresponding
radius of contact, all obtained from the simulation output for the
equilibrium surface.

The gap distance $d$ is varied stepwise and the capillary force is
calculated each time to obtain force-distance curves for a particular
choice of $D$ and $\theta$.

\subsection{Single capillary bridge: Effect of volume}

Surface Evolver simulation results showing the effect of volume on
the maximum capillary force of a single fluid bridge. Since the fluid
is pinned at the top to the same diameter, D, a smaller volume would
result in high interfacial curvatures, which increases the capillary
force due to the negative Laplace pressure. In this case, small contact
angles lead to a greater increase in adhesion.

\begin{figure}[H]

\begin{centering}
\includegraphics{FigureS2-Effect_of_fluid_volume}\caption{Normalised maximum capillary force for a single bridge as a function
of fluid volume}
\par\end{centering}
\end{figure}

\subsection{Capillary Bridge Model: Effect of hair diameter at constant fluid
volume}

Here, instead of scaling the fluid volume relative to the hair diameter,
we now assume a fixed total fluid volume distributed equally among
the $N$ hairs. Total fluid volume, $V_{total}=NV_{f}=2000$. Hair
diameter is varied while keeping the total hair contact area constant.
Length is in arbitrary units. Forces increase at a much smaller rate
on decreasing diameter when compared to the case with self-similar
scaling of fluid volume (Figure \ref{fig:Effect-of-hair}).

\begin{figure}[H]
\includegraphics{FigureS3-Effect_of_hair_size(constant_vol)}

\caption{Normalised adhesion force of hairy pad system on a hydrophilic and
hydrophobic substrate as a function of hair diameter ($D_{h}$), calculated
from the capillary bridge model. The total adhesive fluid volume is
fixed to 2000. Adhesion forces are calculated from minima of the respective
force-distance curves. Negative force value represents attraction.
The bubble's contribution to the net force for an \emph{underwater:
bubble} contact is denoted by plus symbols. Bubble volume and pad
diameter are kept fixed. All lengths are scaled relative to $D_{p}$.}
\end{figure}

\subsection{Statistical comparison}

Pairwise statistical comparison of single leg adhesion force measurements
of the ladybug beetle (\emph{Coccinella septempuctata}) for each contact
type and substrate are shown (Table \ref{tab:Statistical-analysis}).
The uncorrected p-values and Common Language Effect Size (CLES) are
obtained from post-hoc pair-wise Student t-test between A and B while
keeping the third parameter fixed. p-values showing statistically
significant difference between A and B are in boldface. The condition
for statistical significance is based on the Bonferroni-corrected
critical p-value of 0.008.

\begin{table}[H]
\noindent \begin{centering}
\begin{tabular}{|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}|>{\centering}m{0.15\linewidth}|>{\centering}m{0.15\linewidth}|}
\hline 
Fixed & A & B & p-value & CLES\tabularnewline
\hline 
\hline 
In air & PFOTS & Glass & 0.959 & 0.48\tabularnewline
\hline 
Underwater: bubble & PFOTS & Glass & 0.011 & 0.96\tabularnewline
\hline 
Underwater: no bubble & PFOTS & Glass & \textbf{< 0.001} & 1.0\tabularnewline
\hline 
PFOTS & In air & Underwater: bubble & 0.897 & 0.48\tabularnewline
\hline 
PFOTS & In air & Underwater: no bubble & 0.828 & 0.48\tabularnewline
\hline 
PFOTS & Underwater: bubble & Underwater: no bubble & 0.721 & 0.44\tabularnewline
\hline 
Glass & In air & Underwater: bubble & \textbf{0.002} & 1.0\tabularnewline
\hline 
Glass & In air & Underwater: no bubble & \textbf{< 0.001} & 1.0\tabularnewline
\hline 
Glass & Underwater: bubble & Underwater: no bubble & 0.07 & 0.84\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\caption{\label{tab:Statistical-analysis}}
\end{table}

\subsection{Capillary force due to a bubble\label{subsec:Capillary-force-due}}

Capillary force of a single bubble against a PFOTS surface are compared
for two different volumes. The volumes correspond to the expected
range in the case of the trapped bubble in a ladybug. Here, the bubble
is pinned to a micropatterned PDMS substrate on the top. The maximum
adhesion force of any of the bubble never exceeds 50 \textmu N,
significantly lower than the beetle's underwater adhesion to the same
substrate (> 400 \textmu N). Thus, the bubble's contribution
to adhesion in the ``\emph{underwater: bubble}'' contact of a ladybug's
pad should be negligible.

\begin{figure}[H]
\begin{centering}
\includegraphics{FigureS4-Expt_bubble_force}\caption{Capillary force of the bubble}
\par\end{centering}
\end{figure}

>>>>

%% LyX 2.3.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{achemso}
\setkeys{acs}{keywords = true}
\usepackage[T2A,LGR,T1]{fontenc}
\usepackage{babel}
\usepackage{array}
\usepackage{textcomp}
\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subscript}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}
%\usepackage{lineno}
%\linenumbers

@@

\keywords{adhesion; capillary force; underwater;  beetle; gecko}

@@

In air, adhesion forces of the distal pad of ladybug beetle against
glass and PFOTS were similar, i.e. no significant differences were
detected (Figure \ref{fig:Effect-of-contact} and Table S.1).
In contrast, the underwater adhesion on a PFOTS surface was significantly
larger than on glass (p < 0.001). This stronger adhesion on PFOTS
was observed both in the presence and absence of a trapped bubble.
In both cases, the adhesion force reached similar values as in air. In contrast,
on glass, adhesion underwater was significantly reduced when compared
to dry conditions, irrespective of the presence of a trapped bubble
(p $\leq$ 0.002). In the presence of a bubble, underwater adhesion on glass was slightly higher (CLES = 0.84, p = 0.07).

@@

The capillary force, $f$, is the sum of two contributions: surface
tension and Laplace pressure (equation S.1). Force
versus distance for a single capillary bridge was calculated by Surface
Evolver simulations\cite{RN206}  (described in supplementary section S.1),
and used to obtain $F_{net}$ as a function of $d$ for each type
of contact. The adhesion force of the complete hairy pad system was
then obtained from the minima of $F_{net}$, where negative force
values represent attraction.

@@

\subcaption{Force contributions}%
\end{minipage}\caption{\label{fig:Single-bridge}Normalised capillary force of a single liquid
bridge in contact with a substrate and pinned to a circular perimeter
on top. Fluid size parameter, $\phi_{f}=2$. Negative force values
represents attraction. a) Force-distance curves are shown for different
contact angles of the liquid with the substrate. b) Adhesion forces,
calculated from the minima of the corresponding force-distance curves,
are plotted as a function of contact angle with the substrate, together
with its Laplace and surface tension components (equation S.1).
Simulation snapshots of the liquid meniscus corresponding to angles
6\textdegree{} and 150\textdegree{} are depicted.}
\end{figure}

@@

Our experiments demonstrate that the ladybug beetle can attach underwater
to a hydrophobic substrate even without a bubble trapped around its
hairs. A previous study\cite{RN87} proposed that a bubble
is necessary for underwater attachment in beetles. This is, however,
only true for hydrophilic substrates, where a trapped bubble can facilitate
underwater adhesion due to the hairs making contact in a de-wetted
environment. For a hydrophobic substrate, the adhesion is similar
regardless of whether the contact occurs in air or underwater conditions,
with or without a trapped bubble. Our theoretical calculations further
show that the bubble by itself has a negligible capillary contribution
to the net underwater adhesion of the pad. Direct force measurement
of a single similarly sized bubble making contact with a hydrophobic
substrate shows a maximum adhesion less than 50 \textmu N, which
further validates that the bubble's contribution is insignificant
(Figure S.3).

@@

We are grateful to Eduard Arzt and Ren\textcyr{\`\cyre} Hensel for
fruitful discussions. This work was supported by the \emph{Deutsche Forschungsgemeinschaft}
(Grant number: PI 1351/2-1).
